## Question 1

### (b) summary and plot
```{r, summary of the data}
image1 %>%
  group_by(label) %>%
  summarise(count = n()) %>%
  mutate(percentage = count/sum(count))

image2 %>%
  group_by(label) %>%
  summarise(count = n()) %>%
  mutate(percentage = count/sum(count))

image3 %>%
  group_by(label) %>%
  summarise(count = n()) %>%
  mutate(percentage = count/sum(count))

image_binded %>%
  group_by(label) %>%
  summarise(count = n()) %>%
  mutate(percentage = count/sum(count))

image1.valid %>%
  group_by(label) %>%
  summarise(count = n()) %>%
  mutate(percentage = count/sum(count))

image2.valid %>%
  group_by(label) %>%
  summarise(count = n()) %>%
  mutate(percentage = count/sum(count))

image3.valid %>%
  group_by(label) %>%
  summarise(count = n()) %>%
  mutate(percentage = count/sum(count))

image.valid %>%
  group_by(label) %>%
  summarise(count = n()) %>%
  mutate(percentage = count/sum(count))
```

label: -1 (no cloud); 0; 1 (cloud)
image1: 37.3, 28.6, 34.1
image2: 43.8. 38.5, 17.8
image3: 29.3, 52.3, 18.4

```{r, plot}
image1 %>%
  ggplot(aes(x,y)) +
  geom_raster(aes(fill=label)) +
  scale_fill_gradientn(colours = c('grey','black','white'))

image2 %>%
  ggplot(aes(x,y)) +
  geom_raster(aes(fill=label)) +
  scale_fill_gradientn(colours = c('grey','black','white'))

image3 %>%
  ggplot(aes(x,y)) +
  geom_raster(aes(fill=label)) +
  scale_fill_gradientn(colours = c('grey','black','white'))
```

We observe several trends and patterns: (i) Areas of both clouds and surfaces tend to be large. In other words, there is almost no slim slices of clouds and surfaces. This is reflected in the dataset as pixels near to each other tend to have the same labels, unless they are at the boundaries between classes with different labels; (ii) Almost no "black holes" exist in either clouds or surfaces. This pattern shows somewhat continuity of cloud and surface distribution: 
 
The i.i.d. assumption is not justified for the samples. Since we have observed large areas of clouds and surfaces, given a pixel in the image, in most of the scenarios it has the same label as its neighbors, which implies that pixels near each other are highly dependent and tend to have the same label.

### (c) Visual and quantitative EDA

Remark: From Figure 3 in the paper, we could see that the data collected for three different images are consisted of roughly similar components. From the visualization in (b), we could see that the cluster pattern exists in all three images. Since the attributes for these images are similar, it makes sense to combine the attributes when performing EDA analysis.

#### (i)

```{r,cache=TRUE, pairwise relationship}
ggpairs(image_binded,columns = c("NDAI", "SD", "CORR","DF","CF","BF","AF","AN"))
```

We could see some correlation between SD and NDAI (cor = 0.631). The correlation between SD and CORR and that between CORR and NADI are not as significant.

The correlations between NDAI, SD, CORR and the radiance features increase as the angles decrease (from DF to AN) though all the correlations are low (almost all with <0.5 correlations).

For the radiance, we could see that the correlations between those with smaller angles are higher than those with larger angles given similar angle difference. The correlation decreases as the difference between angles increases but all correlations are at least 0.548.

#### (ii)
```{r, relationship between features and labels}
p1 = image_binded %>% mutate(label = as.character(label)) %>%
  ggplot() +
  geom_density(aes(x=SD, fill=label),alpha=0.6)

p2 = image_binded %>% mutate(label = as.character(label),
                        log_SD = log(SD)) %>%
  ggplot() +
  geom_density(aes(x=log_SD, fill=label),alpha=0.6)

p3 = image_binded %>% mutate(label = as.character(label)) %>%
  ggplot() +
  geom_density(aes(x=NDAI, fill=label),alpha=0.6)

p4 = image_binded %>% mutate(label = as.character(label)) %>%
  ggplot() +
  geom_density(aes(x=CORR, fill=label),alpha=0.6)

p5 = image_binded %>% mutate(label = as.character(label)) %>%
  ggplot() +
  geom_density(aes(x=DF, fill=label),alpha=0.6)

p6 = image_binded %>% mutate(label = as.character(label)) %>%
  ggplot() +
  geom_density(aes(x=CF, fill=label),alpha=0.6)

p7 = image_binded %>% mutate(label = as.character(label)) %>%
  ggplot() +
  geom_density(aes(x=BF, fill=label),alpha=0.6)

p8 = image_binded %>% mutate(label = as.character(label)) %>%
  ggplot() +
  geom_density(aes(x=AF, fill=label),alpha=0.6)

p9 = image_binded %>% mutate(label = as.character(label)) %>%
  ggplot() +
  geom_density(aes(x=AN, fill=label),alpha=0.6)

p1+p2+p3+p4

p5+p6+p7+p8+p9
```

For the three essential engineered features (SD,NDAI,CORR): samples labelled "no cloud" typically have lower SD, NDAI and CORR.  

For the five radiance features: samples labelled "no cloud" typically have two peaks, and are more tightly distributed.
Samples labelled "cloud" are more sparsely distributed and negatively skewed. The peak for samples labelled "cloud" normally lies between the peaks for samples labelled "no cloud"

*** Side remark - Also, it appears that almost all data points with low NDAI are labelled as "no cloud", while a few are undetermined and almost none are labelled as "cloud".


```{r, all pairwise relationships}
ggpairs(image_binded,columns = c("NDAI", "SD", "CORR","DF","CF","BF","AF","AN", "label")) 
```